---
title: "Toy Horse Case"
author: "Team 2 - MSBA 2020"
date: "2/9/2020"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(conjoint)
library(cluster)
library(fpc)
library(factoextra)
library(gridExtra)
dir = "C:/Users/liujr/Downloads"
setwd(dir)
load("C:/Users/liujr/Downloads/GBA424 - Toy Horse Case Data.Rdata")


```

The first step is to use the conjoinData to perform individual level regressions and predict missing ratings (NAs).

```{r predict NAs}
#rename the conjoint data
names(conjointData) = c('ID','profile','rating','price_low','height_26','motion_rocking','style_glamorous')

model = list()
for (i in 1:nrow(distinct(conjointData,ID))){
    lm = lm(rating~ price_low+height_26+motion_rocking+style_glamorous,data = conjointData[conjointData$ID==i,])
    model[[i]] = lm$coefficients
}

#the final regression table
individualRegression = as.data.frame(matrix(unlist(model),nrow = 200, byrow=T))
names(individualRegression) = names(model[[1]])
individualRegression = cbind(distinct(conjointData,ID),individualRegression)

#predict the NAs
for(id in 1:200){
    conjointData[conjointData$ID==id,][which(is.na(conjointData[conjointData$ID == id,]$rating)),]$rating = 
        individualRegression[individualRegression$ID==id,]$`(Intercept)`+
        individualRegression[individualRegression$ID==id,]$price_low*conjointData[conjointData$ID==id,][which(is.na(conjointData[conjointData$ID == id,]$rating)),]$price_low+
        individualRegression[individualRegression$ID==id,]$height_26*conjointData[conjointData$ID==id,][which(is.na(conjointData[conjointData$ID == id,]$rating)),]$height_26+
        individualRegression[individualRegression$ID==id,]$motion_rocking*conjointData[conjointData$ID==id,][which(is.na(conjointData[conjointData$ID == id,]$rating)),]$motion_rocking+
        individualRegression[individualRegression$ID==id,]$style_glamorous*conjointData[conjointData$ID==id,][which(is.na(conjointData[conjointData$ID == id,]$rating)),]$style_glamorous    
}

individualRegression$ID = NULL
```

With the predicted ratings and part-utilites for each consumer, we can perform a post-hoc cluster analysis via K-Means clustering. 

First, load the function and then perform the cluster analysis to give optimal number of clusters.

```{r cluster function and optimal clusters}
clustTest = function(toClust,print=TRUE,scale=TRUE,maxClusts=15,seed=12345,nstart=20,iter.max=100){
    if(scale){ toClust = scale(toClust);}
    set.seed(seed);   # set random number seed before doing cluster analysis
    wss <- (nrow(toClust)-1)*sum(apply(toClust,2,var))
    for (i in 2:maxClusts) wss[i] <- sum(kmeans(toClust,centers=i,nstart=nstart,iter.max=iter.max)$withinss)
    ##gpw essentially does the following plot using wss above. 
    #plot(1:maxClusts, wss, type="b", xlab="Number of Clusters",ylab="Within groups sum of squares")
    gpw = fviz_nbclust(toClust,kmeans,method="wss",iter.max=iter.max,nstart=nstart,k.max=maxClusts) #alternative way to get wss elbow chart.
    pm1 = pamk(toClust,scaling=TRUE)
    ## pm1$nc indicates the optimal number of clusters based on 
    ## lowest average silhoutte score (a measure of quality of clustering)
    #alternative way that presents it visually as well.
    gps = fviz_nbclust(toClust,kmeans,method="silhouette",iter.max=iter.max,nstart=nstart,k.max=maxClusts) 
    if(print){
        grid.arrange(gpw,gps, nrow = 1)
    }
    list(wss=wss,pm1=pm1$nc,gpw=gpw,gps=gps)
}
##Runs a set of clusters as kmeans
##Arguments:
##  toClust, data.frame with data to cluster
##  nClusts, vector of number of clusters, each run as separate kmeans 
##  ... some additional arguments to be passed to clusters
##Return:
##  list of 
##    kms, kmeans cluster output with length of nClusts
##    ps, list of plots of the clusters against first 2 principle components
runClusts = function(toClust,nClusts,print=TRUE,maxClusts=15,seed=12345,nstart=20,iter.max=100){
    if(length(nClusts)>4){
        warning("Using only first 4 elements of nClusts.")
    }
    kms=list(); ps=list();
    for(i in 1:4){
        kms[[i]] = kmeans(toClust,nClusts[i],iter.max = iter.max, nstart=nstart)
        ps[[i]] = fviz_cluster(kms[[i]], geom = "point", data = toClust) + ggtitle(paste("k =",nClusts[i]))
        
    }
    library(gridExtra)
    if(print){
        tmp = marrangeGrob(ps, nrow = 2,ncol=2)
        print(tmp)
    }
    list(kms=kms,ps=ps)
}

##Plots a kmeans cluster as three plot report
##  pie chart with membership percentages
##  ellipse plot that indicates cluster definitions against principle components
##  barplot of the cluster means
plotClust = function(km,toClust,discPlot=FALSE){
    nc = length(km$size)
    if(discPlot){par(mfrow=c(2,2))}
    else {par(mfrow=c(3,1))}
    percsize = paste(1:nc," = ",format(km$size/sum(km$size)*100,digits=2),"%",sep="")
    pie(km$size,labels=percsize,col=1:nc)
    
    clusplot(toClust, km$cluster, color=TRUE, shade=TRUE,
             labels=2, lines=0,col.clus=1:nc); #plot clusters against principal components
    
    if(discPlot){
        plotcluster(toClust, km$cluster,col=km$cluster); #plot against discriminant functions ()
    }
    rng = range(km$centers)
    dist = rng[2]-rng[1]
    locs = km$centers+.05*dist*ifelse(km$centers>0,1,-1)
    bm = barplot(km$centers,beside=TRUE,col=1:nc,main="Cluster Means",ylim=rng+dist*c(-.1,.1))
    text(bm,locs,formatC(km$centers,format="f",digits=1))
}

Checks=clustTest(individualRegression,print=TRUE,scale=TRUE,maxClusts=15,seed=12345,nstart=20,iter.max=100)
clusts=runClusts(individualRegression,c(2,3,4,5),print=TRUE,maxClusts=15,seed=12345,nstart=20,iter.max=100)
plotClust(clusts[[1]][[2]],individualRegression)

```

Now to actually perform the K-Means cluster analysis and the output plot will show the optimal clusters. 

```{r run k-means cluster analysis}
#Run K-means cluster analysis
clusts=runClusts(individualRegression,c(2,3,4,5),print=TRUE,maxClusts=15,seed=12345,nstart=20,iter.max=100)
```

Now plot our clusters to look at the attributes of each.

```{r plot k-means cluster analysis}
#Run K-means cluster analysis
plotClust(clusts[[1]][[2]],individualRegression)
```

Now we add demograhic data to the conjoin data

```{r add demographic data to conjoint data}
#Join demographic data onto conjointData
merge_df = merge(conjointData, respondentData, by = 'ID', all.x = TRUE)

#test whether the priori segments are significant 
significanceAge = lm(rating~(price_low+height_26+motion_rocking+style_glamorous)*age,data = merge_df)
summary(significanceAge)
#age will affect the part utilities of height and motion

significanceGender = lm(rating~(price_low+height_26+motion_rocking+style_glamorous)*gender,data = merge_df)
summary(significanceGender)
#gender will affect the part utilities of all attribute

lm_female = lm(rating~.-ID-profile-gender-age, data = merge_df[merge_df$gender == 1,])
lm_male = lm(rating~.-ID-profile-gender-age, data = merge_df[merge_df$gender == 0,])
par(mfrow = c(1, 2))
barplot(lm_female$coefficients[-1])
barplot(lm_male$coefficients[-1])
# felame's preference is as the same as cluster3's , 
# male's preference is as the same as cluster1's

lm_34 = lm(rating~.-ID-profile-gender-age, data = merge_df[merge_df$age == 1,])
lm_2 = lm(rating~.-ID-profile-gender-age, data = merge_df[merge_df$age == 0,])
par(mfrow = c(1, 2))
barplot(lm_34$coefficients[-1])
barplot(lm_2$coefficients[-1])
#other new products can be designed to match different age groups' preference

```

Merge cluster data and load financial/marketshare data
```{r cluster and financial data}
#join cluster data into individual regression
clusts$kms[[2]][1]
individualRegression$segmentation = unlist(clusts$kms[[2]][1])
individualRegression$gender = NA
for (i in 1:200){
    individualRegression[i,]$gender = respondentData[i,]$gender
}
table(individualRegression$gender,individualRegression$segmentation)
# can explain why female and cluster3 have the same preference, and male and #cluster1 have the same preference
#    1  2  3
# 0 49 29 14
# 1 31 24 53
# female - cluster3, because most of them are female
# male - cluster1, beacuase most of them are male
# so we can definitely launch product1 and product3

# Round conjointData - Ratings to nearest 5
mround = function(x,base){
    base*round(x/base)}

conjointData$rating = mround(conjointData$rating, 5)

#Complete Ratings
library("reshape")
cast_rating = cast(conjointData[1:3],ID~profile)
cast_rating = cast_rating[-1]


simFCDecisions = function(scen,data){ 
    inmkt = data[,scen] #construct the subsetted matrix of options
    len = length(scen)
    inmkt$bestOpts = apply(inmkt,1,max)
    for (i in 1:200){
        inmkt[i,1:len] = (inmkt[i,1:len] == inmkt$bestOpts[i])
        inmkt[i,1:len] = inmkt[i,1:len]/sum(inmkt[i,1:len])
    }
    
    ret = inmkt[1:len]
    names(ret) = names(inmkt[1:len])
    ret #decisions
}

calcUnitShares = function(decisions){
    round(colSums(decisions)/sum(decisions),2) #assumes that total decisions is market size
}

simFCShares=function(scen,data){
    decs = simFCDecisions(scen,data) #determine decisions
    calcUnitShares(decs) #calculate shares and return
}

simFCScenarios = function(scenarios,data,...){
    res = matrix(nrow=length(scenarios),ncol=length(data)) #sets everything to NA by default
    for(i in 1:length(scenarios)){ ##loop over scenarios
        res[i, scenarios[[i]] ] = simFCShares(scenarios[[i]],data,...)##  calculate market shares and save to right columns in res for the scenario
    }
    res = as.data.frame(res); names(res) = names(data)
    res ##return result table
}

#Design new products according to different segmentaions' preference
#p1 for cluster1
p1 = unique(conjointData$profile[conjointData$price_low==1&conjointData$height_26==1&conjointData$motion_rocking==0&conjointData$style_glamorous==0])
names(cast_rating)[p1] = 'p1'
#p2E for cluster2(exsiting)
p2E = unique(conjointData$profile[conjointData$price_low==0&conjointData$height_26==0&conjointData$motion_rocking==1&conjointData$style_glamorous==1])
names(cast_rating)[p2E] = 'p2E'
#p3 for cluster3
p3 = unique(conjointData$profile[conjointData$price_low==1&conjointData$height_26==1&conjointData$motion_rocking==1&conjointData$style_glamorous==1])
names(cast_rating)[p3] = 'p3'
#current exsiting product
pE = unique(conjointData$profile[conjointData$price_low==0&conjointData$height_26==0&conjointData$motion_rocking==1&conjointData$style_glamorous==0])
names(cast_rating)[pE] = 'pE'
#p4 for age 3-4
p4 = unique(conjointData$profile[conjointData$price_low==1&conjointData$height_26==1&conjointData$motion_rocking==0&conjointData$style_glamorous==1])
names(cast_rating)[p4] = 'p4'
#p2L try to lower the price for cluster2(lower price version p2E)
p2L = unique(conjointData$profile[conjointData$price_low==1&conjointData$height_26==0&conjointData$motion_rocking==1&conjointData$style_glamorous==1])
names(cast_rating)[p2L] = 'p2L'

#c1 competitor's current product
c1 = unique(conjointData$profile[conjointData$price_low==0&conjointData$height_26==1&conjointData$motion_rocking==1&conjointData$style_glamorous==0])
names(cast_rating)[c1] = 'c1'
#c2 competitor might lower price
c2 = unique(conjointData$profile[conjointData$price_low==1&conjointData$height_26==1&conjointData$motion_rocking==1&conjointData$style_glamorous==0])
names(cast_rating)[c2] = 'c2'

#load finance data and join it into profiles data
financeData = data.frame(product = c('profile', 'VC','price','margin'), p1 = c(4,29,96,67),
           p2E = c(13,33,112,79),
           p3 = c(15,41,112,71),
           pE = c(5,33,112,79),
           p4 = c(12,29,96,67),
           p5 = c(14,33,96,63),
           c1 = c(7,41,112,71),
           c2.low = c(8,41,96,55)
           )
selectNum = c(p1,p2E,p3,pE,p4,c1,c2)
profilesData$margin = NA
profilesData[selectNum,]$margin = unlist(financeData[4,c(2,3,4,5,6,8,9)])
profilesData[p2L,]$margin = unlist(financeData[4,3])
```

Now we check to see competitions decisions:

In current situation:
```{r will competitor lower price in current situation}
##################################################################################
#we need to judge if our competitor will lower its price
#simulation:In current situation(c1,p2E,pE),if the competitor lower its #price(c1->c2)
scens1 = list()
scens1[[1]]=c(c1,p2E,pE)
scens1[[2]]=c(c2,p2E,pE)
simFCScenarios(scens1,cast_rating)
df1 = as.data.frame(simFCScenarios(scens1,cast_rating))
df1$mkshare = NA
df1[1,]$mkshare = rowSums(df1[1,-c1],na.rm = T)
df1[2,]$mkshare = rowSums(df1[2,-c2],na.rm = T)
#calculate our margin
myProduct = c(p1,p2E,p3,pE,p4,p2L)
df1$margin = NA
for(r in 1:nrow(df1)){
    df1[r,]$margin = sum(profilesData[myProduct ,]$margin*df1[r,myProduct]*4000,na.rm = T)
    
}

#calculate the FC and profit for 1st year 
#Costs = 20,000/year * #products + $20,000/3 *#products not in existing set
newProduct = c(p1,p3,p4,p2L)
df1$firstYearFC = NA
for (r in 1:nrow(df1)){
    df1[r,]$firstYearFC = sum(sum(!is.na(df1[r,myProduct]))*20000)+sum(sum(!is.na(df1[r,newProduct]))*6667)
}
df1$firstYearProfit = df1$margin-df1$firstYearFC

#profit after first year
df1$laterYearFC = NA
for (r in 1:nrow(df1)){
    df1[r,]$laterYearFC = sum(sum(!is.na(df1[r,myProduct]))*20000)
}
df1$laterYearProfit = df1$margin-df1$laterYearFC

#calculate the profit for competitor
c1
df1$comMargin = NA
df1[1,]$comMargin = sum(profilesData[c1 ,]$margin*df1[1,c1]*4000,na.rm = T)
c2
df1[2,]$comMargin = sum(profilesData[c2 ,]$margin*df1[2,c2]*4000,na.rm = T)
    
df1$comCurrentFC = 20000

df1$comCurrentProfit = df1$comMargin-df1$comCurrentFC
df1
#we found that the competitor's profit will increase 
#so in current situation(c1,p2E,pE),our competitor will lower its price(switch c1 to c2)
#write.csv(df1,'c1toc2InCurrentMk.csv')
```
The competitor should lower his price in the curretn sitation.


Check optimal product combination with competitors current price.
```{r Check profitability of products with c1}
#what if our competitor will take actions, after we launching the best products?
scens2 = list()
scens2[[1]]=c(c1,p2E,pE) 
scens2[[2]]=c(c1,p2E) 
scens2[[3]]=c(c1,pE)
#we can find that p2E performed better(higher market share), so consider to drop pE
scens2[[4]]=c(c1,p2E,p1)    
scens2[[5]]=c(c1,p2E,p3)    
scens2[[6]]=c(c1,p2E,p1,p3) 
scens2[[7]]=c(c1,p1,p3)     
scens2[[8]]=c(c1,p2E,p4)     
scens2[[9]]=c(c1,p2E,p2L)     
scens2[[10]]=c(c1,p2E,p4,p2L) 
scens2[[11]]=c(c1,p1,p4)
scens2[[12]]=c(c1,p1,p2L)     
scens2[[13]]=c(c1,p3,p4)     
scens2[[14]]=c(c1,p3,p2L)
scens2[[15]]=c(c1,p2E,p1,p4)
scens2[[16]]=c(c1,p2E,p1,p2L)
scens2[[17]]=c(c1,p2E,p3,p4)
scens2[[18]]=c(c1,p1,p3,p4)
scens2[[19]]=c(c1,p1,p3,p2L)
scens2[[20]]=c(c1,p3,p4,p2L)
df2 = as.data.frame(simFCScenarios(scens2,cast_rating))
df2$mkshare = rowSums(df2[-c1],na.rm = T)

#calculate our margin
myProduct = c(p1,p2E,p3,pE,p4,p2L)
df2$margin = NA
for(r in 1:nrow(df2)){
    df2[r,]$margin = sum(profilesData[myProduct ,]$margin*df2[r,myProduct]*4000,na.rm = T)
    
}

#calculate the FC and profit for 1st year 
#Costs = 20,000/year * #products + $20,000/3 *#products not in existing set
newProduct = c(p1,p3,p4) #p2L is not a new product because we just lower the price of p2E to get it 
df2$firstYearFC = NA
for (r in 1:nrow(df2)){
    df2[r,]$firstYearFC = sum(sum(!is.na(df2[r,myProduct]))*20000)+sum(sum(!is.na(df2[r,newProduct]))*6667)
}
df2$firstYearProfit = df2$margin-df2$firstYearFC

#profit after first year
df2$laterYearFC = NA
for (r in 1:nrow(df2)){
    df2[r,]$laterYearFC = sum(sum(!is.na(df2[r,myProduct]))*20000)
}
df2$laterYearProfit = df2$margin-df2$laterYearFC

#calculate the profit for competitor
c1
df2$comMargin = NA
for(r in 1:nrow(df2)){
    df2[r,]$comMargin = sum(profilesData[c1 ,]$margin*df2[r,c1]*4000,na.rm = T)
    
}
df2$comCurrentFC = NA
for (r in 1:nrow(df2)){
    df2[r,]$comCurrentFC = sum(sum(!is.na(df2[r,c1]))*20000)
}
df2$comCurrentProfit = df2$comMargin-df2$comCurrentFC
df2
#write.csv(df2,'currentMk.csv')
#we found that in current market(competitor sells c1),launching p2L and p3 can get the highest profit
```

Launching p2L and p3 is most profitable if competitors does not lower his price. (But he likely will.)


Now check competitors profit in scenarios given he lowers his price.
```{r }
scens3 = list()
scens3[[1]]=c(c2,p2L,p3)       
df3 = as.data.frame(simFCScenarios(scens3,cast_rating))
df3$mkshare = rowSums(df3[-c2],na.rm = T)

#calculate our margin
myProduct = c(p1,p2E,p3,pE,p4,p2L)
df3$margin = NA
for(r in 1:nrow(df3)){
    df3[r,]$margin = sum(profilesData[myProduct ,]$margin*df3[r,myProduct]*4000,na.rm = T)
    
}

#calculate the FC and profit for 1st year 
#Costs = 20,000/year * #products + $20,000/3 *#products not in existing set
newProduct = c(p1,p3,p4)
df3$firstYearFC = NA
for (r in 1:nrow(df3)){
    df3[r,]$firstYearFC = sum(sum(!is.na(df3[r,myProduct]))*20000)+sum(sum(!is.na(df3[r,newProduct]))*6667)
}
df3$firstYearProfit = df3$margin-df3$firstYearFC

#profit after first year
df3$laterYearFC = NA
for (r in 1:nrow(df3)){
    df3[r,]$laterYearFC = sum(sum(!is.na(df3[r,myProduct]))*20000)
}
df3$laterYearProfit = df3$margin-df3$laterYearFC

#calculate the profit for competitor
c2
df3$comMargin = NA
for(r in 1:nrow(df3)){
    df3[r,]$comMargin = sum(profilesData[c2 ,]$margin*df3[r,c2]*4000,na.rm = T)
    
}

df3$comCurrentFC = NA
for (r in 1:nrow(df3)){
    df3[r,]$comCurrentFC = sum(sum(!is.na(df3[r,c2]))*20000)
}

df3$comCurrentProfit = df3$comMargin-df3$comCurrentFC
df3
# write.csv(df3,'afterP2LP3Launch.csv')
#obviously if we launch p2L and p3, then lowering price(changing c1 to c2) can increase competitor's profit
#so we can tell that in whatever situation, our competitor must lower its price
#and we can just focous on the situation that our competitor launch c2(figure out the optimal product combinations when competitor launch c2)

```

The competitors makes a higher profit with a lower price in every scenario, so we can conclude he will likely lower his price.

Now we can run market share/ profitability scenarios assuming the competitor lowers his price.
``` {r optimal product offerings}
######market simulation when competitor sells c2########
scens4 = list()
scens4[[1]]=c(c2,p2E,pE) 
scens4[[2]]=c(c2,p2E) 
scens4[[3]]=c(c2,pE)
#we can find that p2E performed better(higher mk share), so we consider to drop pE
scens4[[4]]=c(c2,p2E,p1)    
scens4[[5]]=c(c2,p2E,p3)    
scens4[[6]]=c(c2,p2E,p1,p3) 
scens4[[7]]=c(c2,p1,p3)     
scens4[[8]]=c(c2,p2E,p4)     
scens4[[9]]=c(c2,p2E,p2L)     
scens4[[10]]=c(c2,p2E,p4,p2L) 
scens4[[11]]=c(c2,p1,p4)
scens4[[12]]=c(c2,p1,p2L)     
scens4[[13]]=c(c2,p3,p4)     
scens4[[14]]=c(c2,p3,p2L)
scens4[[15]]=c(c2,p2E,p1,p4)
scens4[[16]]=c(c2,p2E,p1,p2L)
scens4[[17]]=c(c2,p2E,p3,p4)
scens4[[18]]=c(c2,p1,p3,p4)
scens4[[19]]=c(c2,p1,p3,p2L)
scens4[[20]]=c(c2,p3,p4,p2L)
df4 = as.data.frame(simFCScenarios(scens4,cast_rating))
df4$mkshare = rowSums(df4[-c2],na.rm = T)
#calculate our margin
myProduct = c(p1,p2E,p3,pE,p4,p2L)
df4$margin = NA
for(r in 1:nrow(df4)){
    df4[r,]$margin = sum(profilesData[myProduct ,]$margin*df4[r,myProduct]*4000,na.rm = T)
    
}

#calculate the FC and profit for 1st year 
#Costs = 20,000/year * #products + $20,000/3 *#products not in existing set
newProduct = c(p1,p3,p4)
df4$firstYearFC = NA
for (r in 1:nrow(df4)){
    df4[r,]$firstYearFC = sum(sum(!is.na(df4[r,myProduct]))*20000)+sum(sum(!is.na(df4[r,newProduct]))*6667)
}
df4$firstYearProfit = df4$margin-df4$firstYearFC

#profit after first year
df4$laterYearFC = NA
for (r in 1:nrow(df4)){
    df4[r,]$laterYearFC = sum(sum(!is.na(df4[r,myProduct]))*20000)
}
df4$laterYearProfit = df4$margin-df4$laterYearFC

#calculate the profit for competitor
c2
df4$comMargin = NA
for(r in 1:nrow(df4)){
    df4[r,]$comMargin = sum(profilesData[c2 ,]$margin*df4[r,c2]*4000,na.rm = T)
    
}

df4$comCurrentFC = NA
for (r in 1:nrow(df4)){
    df4[r,]$comCurrentFC = sum(sum(!is.na(df4[r,c2]))*20000)
}

df4$comCurrentProfit = df4$comMargin-df4$comCurrentFC
df4
#write.csv(df4,'finalResult.csv')
#if the competitor launch c2, we should launch p1, p2L, and p3 to generate the highest profits

```

Finally, we see the highest estimated profit is from offering 3 new products (dropping our current selection).

These products are p1,p3, and p2L
p1: Low Price, 26 inches, bouncing, and racing
p2: Low price, 26 inches, rocking, and glamorous
p2L: Low price, 18inches, rocking, and glamorous
